---
title: "lab9 part 2 - chi square with lizards"
author: "Lucas Boyd"
date: "11/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)


```
## read in data
```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```

Get counts of lizards by species

```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n) # arrange data in decreasing order by n
```

## make a subset

-  only UTST lizards
-  Sites: IBPE, RABB, EAST
-  tail condition is recorded (W or B)

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST") ) %>% 
  filter(tail %in% c("B", "W"))
  
```

## Find counts and proportions of lizards with broken or whole tails at each site
```{r}
utst %>% 
  count(site, tail)
```
Use tabyl() to make a nice clean table of data for chisquare test.
```{r}
tail_counts <- utst %>% 
  tabyl(site, tail)

# turn those counts into percentages (default is calculated by rows)
tail_proportions <- tail_counts %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting(digits = 2)  %>% # make the percentages a nice little two digits
  adorn_ns() # add the actual counts to the table
```
## Convert to a contingency table

Convert columns into row names using `column_to_rowname()`
```{r}
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site") # turn the site column into a row name
```

## chi-square
Are tail breakage and site independent?
```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

lizard_tidy <- lizard_x2 %>% # make a nice table of the chisq output to make it easier for inline referencing
  tidy(lizard_x2)

```
Reject the null hypothesis of independence. 
There is a significant association between site and tail breakage. 

The is a significant association between site and lizard tail breakage ($\chi$^2^(`r (lizard_tidy$parameter)`) = `r round(lizard_tidy$statistic, 2)`, *p* = `r round(lizard_tidy$p.value, 3)`).




